---
title: "PA1_template"
author: "Wichan"
date: "8/29/2021"
output:
  html_document: default
---
# Loading and preprocessing the data

##Show any code that is needed to

1. Load the data (i.e.read.csv())

2. Process/transform the data (if necessary) into a format suitable for your analysis

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# libraries
library(ggplot2)
library(dplyr)
# Create a new object by import .csv file
activity <- read.csv("activity.csv")
```


```{r, echo=TRUE}
# look at the structure of activity
str(activity)
```

There are 3 variables in the dataset:

1. steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)

2. date: The date on which the measurement was taken in YYYY-MM-DD format

3. interval: Identifier for the 5-minute interval in which measurement was taken

# Total number of steps taken per day
##For this part of the assignment, you can ignore the missing values in the dataset.

## 1. Calculate the total number of steps taken per day

```{r, echo=TRUE}
StepsPerDay <- aggregate(activity$steps, list(activity$date), FUN=sum)
colnames(StepsPerDay) <- c("Date", "Steps")
StepsPerDay
```

## 2. Make a histogram of the total number of steps taken each day

```{r histogram1, echo=TRUE, fig.width=10, warning=FALSE}
g <- ggplot(StepsPerDay, aes(Steps))
g+geom_histogram(boundary=0, binwidth=2500, col="steelblue", fill="steelblue4")+
  ggtitle("Histogram of the total steps per day")+
  xlab("Steps")+ylab("Frequency")+
  theme(plot.title = element_text(face="bold", size=12))+
  scale_x_continuous(breaks=seq(0,25000,2500))+
  scale_y_continuous(breaks=seq(0,18,2))
```

## 3. Calculate and report the mean and median total number of steps taken per day

```{r, echo=TRUE}
# Mean
mean(StepsPerDay$Steps, na.rm=TRUE)
# Median
median(StepsPerDay$Steps, na.rm=TRUE)
```

# Average daily activity pattern?

## 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r, timeplot1, echo=TRUE, fig.width=10, warning=FALSE}
# create table with steps per time
StepsPerTime <- aggregate(steps~interval,data=activity,FUN=mean,na.action=na.omit)
# variable time (more comprensible for the graph axis)
StepsPerTime$time <- StepsPerTime$interval/100
# Create the line plot
h <- ggplot(StepsPerTime, aes(time, steps))
h+geom_line(col="navyblue")+
  ggtitle("Average steps per time interval")+
  xlab("Time")+ylab("Steps")+
  theme(plot.title = element_text(face="bold", size=12))
```

## 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, echo=TRUE, fig.width=10, warning=FALSE}
# table for dplyr
ST <- tbl_df(StepsPerTime)
# find the column
ST %>% select(time, steps) %>% filter(steps==max(ST$steps))
```


# Imputing missing values

##Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

## 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r, echo=TRUE}
# table for dplyr
activity_with_na <- tbl_df(activity)
# find the column
activity_with_na %>% filter(is.na(steps)) %>% summarize(missing_values = n())
```

## 2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

```{r, echo=TRUE}
# fill in all of the missing values (NAs)
activity$CompleteSteps <- ifelse(is.na(activity$steps), round(StepsPerTime$steps[match(activity$interval, StepsPerTime$interval)],0), activity$steps)
```

## 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, echo=TRUE}
# Create a new dataframe "activity_replace_na"
activity_replace_na <- data.frame(steps=activity$CompleteSteps, interval=activity$interval, date=activity$date)
# show the first 10 rows of activity_replace_na
head(activity_replace_na, n=10)
```

## 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r histogram2, echo=TRUE, fig.width=10, warning=FALSE}
# prepare data
StepsPerDay_ReplaceNA <- aggregate(activity_replace_na$steps, list(activity_replace_na$date), FUN=sum)
colnames(StepsPerDay_ReplaceNA) <- c("Date", "Steps")
# Create the histogram
g <- ggplot(StepsPerDay_ReplaceNA, aes(Steps))
g+geom_histogram(boundary=0, binwidth=2500, col="steelblue", fill="steelblue4")+
  ggtitle("Histogram of steps per day")+
  xlab("Steps")+ylab("Frequency")+
  theme(plot.title = element_text(face="bold", size=12))+
  scale_x_continuous(breaks=seq(0,25000,2500))+
  scale_y_continuous(breaks=seq(0,26,2))
```

### Calculate and report the mean and median total number of steps taken per day

```{r, echo=TRUE}
# Mean
mean(StepsPerDay_ReplaceNA$Steps)
#Median
median(StepsPerDay_ReplaceNA$Steps)
```


#Are there differences in activity patterns between weekdays and weekends?

## For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

## 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r echo=TRUE}
# Create variable with date in correct format
activity_replace_na$RealDate <- as.Date(activity_replace_na$date, format = "%Y-%m-%d")
# create a variable with weekdays name
activity_replace_na$weekday <- weekdays(activity_replace_na$RealDate)
# create a new variable indicating weekday or weekend
activity_replace_na$DayType <- ifelse(activity_replace_na$weekday=='Saturday' | activity_replace_na$weekday=='Sunday', 'weekend','weekday')
# show the first 10 rows of activity_replace_na
head(activity_replace_na, n=10)
```

## 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r timeplot2, echo=TRUE, fig.width=10, warning=FALSE}
# Create table with steps per time by weekday and weekend days
StepsPerTimeDT <- aggregate(steps~interval+DayType,data=activity_replace_na,FUN=mean,na.action=na.omit)
# Create variable time
StepsPerTimeDT$time <- StepsPerTime$interval/100
# Create line plot
j <- ggplot(StepsPerTimeDT, aes(time, steps))
j+geom_line(col="steelblue")+
  ggtitle("Comparing average steps per time between weekdays and weekends")+
  xlab("Time")+ylab("Steps")+theme(plot.title = element_text(face="bold", size=12))+
  facet_grid(DayType ~ .)
```
